Parity-odd Neutrino Torque Detection

### 2018-12-29 reply_to_editor+referee_v2.1.txt

Reply to the editor

Dear Editor,

Thank you for your consideration of our manuscript and thank the referee to give valuable comments. We think we can can address all the questions. Especially for the major comment, the referee should look at section 4.2 of 1804.07306 and their conclusion that spin 'direction' correlation is strong. Fig.2 of 1804.07306 shows the uncorrelated spin 'amplitude' correlation, which is unrelated to our work. We also explained it to the referee below.

We think our manuscript appropriate for PRL is that we proposed a new approach and observable to measure the neutrino mass, which has not been considered before. Most proposed neutrino observables are not proven to work, including weak lensing. It is too early to prove that this method works, but rather it is meant to encourage further studies in this new field of angular momentum directions.

Sincerely,

Authors

Reply to the referee

Dear referee,

Thank you for your comments and suggestions in our paper. We answer your comments below and marked the modifications in our paper.

We are proposing a totally new approach to measure neutrino mass, and we have proven that there is a new observable that has not been considered before. Most proposed neutrino observables are not proven to work, not even CMB lensing because of the baryon feedback. The importance of our paper is that it is completely new and is meant to encourage new research studies on the angular momentum connection.

----------------------------------------------------------------------
Report of Referee A -- LX16362/Yu
----------------------------------------------------------------------

Major comments:

My biggest concern with this work is that I think the authors need to
assume that the angular momentum of galaxies traces the angular
momentum of halos. I have done a quick search in the literature and
found 1804.07306, where in Fig. 2 I can see a large scatter between
the spin of galaxies and the spin of their host halos. The authors
need to address this point in detail and explain how they will
distinguish the expected small neutrino angular momentum contribution
from the underlying scatter.

>> Answer:
We agree that the spin connection between galaxies and halos is an important concern. Most literature study the alignment of shapes, or only the amplitude of the angular momentum. Most parts of 1804.07306 focus on the ‘amplitude’ of the angular momentum.
In 1804.07306, is says “It is intriguing that the amplitudes of spins are uncorrelated although the alignment of the spin vectors is relatively good”. In Fig. 2 of 1804.07306, it shows the correlation between angular momentum ‘amplitudes’, not the correlation between ‘directions’. In 1804.07306, Section 4.2, 5.2 and Figure 4, 7 discussed the spin ‘direction’ correlations, and their most important conclusion is that “the median cos\theta_{gas,halo} is 0.71 to 0.73”. This suggests a lower limit of the spin direction correlation between DM halos and baryons, since in higher resolution simulations the angular momentum direction is better conserved. Although 1804.07306 did not study the Eulerian-Lagrangian spin direction correlation, from results from our work it indicates that stars and gas correlates should also correlate with Lagrangian spin direction. The importance of our paper means to start new research studies on the angular momentum connection.
We modified the last part of the introduction clearer — “Recent hydrodynamic galaxy formation simulations suggest that, although uncorrelated in amplitude, the spin directions of stellar, cold gas components and CDM halo are strongly correlated (e.g. section 4.2 of \citep{2018arXiv180407306J}).”

Intermediate comments:

The authors argue that the spin of a galaxy changes very modestly with
distance, so they relate halo spin to galaxy spin. However, this can
not be true in galaxies clusters, where different galaxies inside the
cluster will have different angular momentum. The authors should
explain how this measurement will be done in practice. They should
also explain how they will deal with satellites galaxies.

>> Answer:
Thanks for this comment. 1804.07306 uses only zoom-in simulations and lower mass halos. In ref \citep{2016MNRAS.460.3772S} they use cosmological simulations and found that the central galaxy, DM halo and the distribution of satellite galaxies are aligned with their minor axis of their moments of inertia. The minor axis of the moment of inertia is an indicator of spin axis, however this requires further studies. Related to our work, in clusters most galaxies are elliptical which we are not considering. Only 1% of galaxies are in clusters, so it is not a major effect.
In the first paragraph of discussion, we added “We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.”
Our paper is not trying to prove the applicability, but to propose that there is a new neutrino observable.  And we have proven that there is a new observable that has not been considered before. Most proposed neutrino observables are not ‘proven’ to work, not even CMB lensing (also because of baryon feedback).  The importance of the paper is that it is completely new, and is meant to start new research studies on the angular momentum connection.

The authors state that “A decayed tidal field is hard to torque small
and round objects, so the spin are expected to be contributed mostly
in the linear regime”. While I agree with this statement, I think the
authors can not neglect the contribution of non-linear and baryonic
effects, that although are expected to be small, so is the signal from
neutrinos. The authors write that CDM and baryons are torqued by the
same gravitational shear, such as the spin direction is not affected
by baryonic effects. I agree with the authors for halos, but not for
individual galaxies, where AGN feedback, dynamic friction…etc can
potentially affect the galaxy properties.

>> Answer:
We agree and try to describe in the text more clearly.
Firstly, “A decayed tidal field is hard to torque small and round objects, so the spin are expected to be contributed mostly in the linear regime” is to physically and qualitatively explain why the halo spin between Lagrangian space, Eulerian space and tidal torque theory is as high as 0.8, and we rewrote the CDM-baryon relations in the next paragraph separately, as —
“In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so their subsequent local evolution should not systematically change their respective spin directions. For example, baryonic feedback has to conserve angular momentum, unless the baryons are expelled from the galaxy and carry angular momentum away systematically. The dynamical friction between baryons and dark matter can only increase the angular momentum correlation. These all indicate a high correlation between CDM and baryons (e.g. \citep{2018arXiv180407306J}).”

I agree with the authors that neutrinos should contribute to the total
angular momentum of a halo by a predictable amount. However, I think
the Lagrangian region enclosing the neutrinos that end up in the halo
will be very large, due to the neutrinos thermal velocities. How do
the authors deal with that?

>> Answer:
Yes. We agree. Since initially only neutrinos have large velocity dispersion, not CDM and baryons, so the Lagrangian region for neutrinos (enclosed in the final halo region) is very large, while CDM and baryons should be equivalent. Note that the neutrinos enclosed in the halo region are not necessarily gravitationally bound and they do not evolve with CDM and baryons, so their spin (indeed very small) should not contribute to the CDM halo and baryons. Even considering it, the neutrino faction is small, and the spin is small, so it is a second order effect. The effect neutrino tides torques the CDM+baryon inertia is what we consider.
We rewrote the sentence as — “Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to CDM and baryons.”

Small comments:

Introduction: update neutrino constraints from Planck 2018

>> Answer: Thanks. We updated it.

When the authors mention: “In the cosmic structure formation,
neutrinos are non-relativistic”, they should state that this is a low
redshift, not at high-redshift, were neutrinos are relativistic.

>> Answer: Thanks. We modify the sentence as — “At low redshifts, neutrinos become non-relativistic, and contribute to…”

I do not agree with the authors that a clean observable is the effect
of neutrinos on the weak lensing power spectrum. There are a large set
of systematics that affect these observations, and from the theory
side it is still unclear the contamination by baryonic effects.

>> Answer: 
Weak lensing effect is purely gravitational. We agree with you that it is not clean in the systematics and baryonic effects. We modify the sentence as — “Thus most popular cosmological neutrino mass measurements are not proven to work. For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.”

>> Thank you again for all the comments and suggestions. All the modified parts in the paper are marked with blue text.

Authors

### 2018-12-29 reply_to_referee_v2.0.txt

>> Answer:
We agree that the spin connection between galaxies and halos is an important concern. Most literature study the alignment of shapes, or only the amplitude of the angular momentum. Most parts of 1804.07306 focus on the ‘amplitude’ of the angular momentum.
In 1804.07306, is says “It is intriguing that the amplitudes of spins are uncorrelated although the alignment of the spin vectors is relatively good”. In Fig. 2 of 1804.07306, it shows the correlation between angular momentum ‘amplitudes’, not the correlation between ‘directions’. In 1804.07306, Section 4.2, 5.2 and Figure 4, 7 discussed the spin ‘direction’ correlations, and their most important conclusion is that “the median cos\theta_{gas,halo} is 0.71 to 0.73”. This suggests a lower limit of the spin direction correlation between DM halos and baryons, since in higher resolution simulations the angular momentum direction is better conserved. Although 1804.07306 did not study the Eulerian-Lagrangian spin direction correlation, from results from our work it indicates that stars and gas correlates should also correlate with Lagrangian spin direction. The importance of our paper means to start new research studies on the angular momentum connection.
We modified the last part of the introduction clearer — “Recent hydrodynamic galaxy formation simulations suggest that, although uncorrelated in amplitude, the spin directions of stellar, cold gas components and CDM halo are strongly correlated (e.g. section 4.2 of \citep{2018arXiv180407306J}).”

Intermediate comments:

The authors argue that the spin of a galaxy changes very modestly with
distance, so they relate halo spin to galaxy spin. However, this can
not be true in galaxies clusters, where different galaxies inside the
cluster will have different angular momentum. The authors should
explain how this measurement will be done in practice. They should
also explain how they will deal with satellites galaxies.

>> Answer:
Thanks for this comment. 1804.07306 uses only zoom-in simulations and lower mass halos. In ref \citep{2016MNRAS.460.3772S} they use cosmological simulations and found that the central galaxy, DM halo and the distribution of satellite galaxies are aligned with their minor axis of their moments of inertia. The minor axis of the moment of inertia is an indicator of spin axis, however this requires further studies. Related to our work, in clusters most galaxies are elliptical which we are not considering. Only 1% of galaxies are in clusters, so it is not a major effect.
In the last part of the introduction we added “ In galaxy clusters, central galaxies and satellites are also found to be aligned with halos \citep{2016MNRAS.460.3772S}.”
And, in the first paragraph of discussion, we added “For galaxy clusters, the correlation between minor axes in \cite{2016MNRAS.460.3772S} also implies the correlation between spins. After all, however, since most galaxies are not in clusters so the contribution of clusters is not a major effect.”
Our paper is not trying to prove the applicability, but to propose that there is a new neutrino observable.  And we have proven that there is a new observable that has not been considered before. Most proposed neutrino observables are not ‘proven’ to work, not even CMB lensing (also because of baryon feedback).  The importance of the paper is that it is completely new, and is meant to start new research studies on the angular momentum connection.

@@

When the authors mention: “In the cosmic structure formation,
neutrinos are non-relativistic”, they should state that this is a low
redshift, not at high-redshift, were neutrinos are relativistic.

>> Answer: Thanks. We modify the sentence as — “At low redshifts, neutrinos become non-relativistic, and contribute to…”

I do not agree with the authors that a clean observable is the effect
of neutrinos on the weak lensing power spectrum. There are a large set
of systematics that affect these observations, and from the theory
side it is still unclear the contamination by baryonic effects.

>> Answer: We agree with it. We modify the sentence as — “For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.”

>> Thank you again for all the comments and suggestions. All the modified parts in the paper are marked with blue text.
Authors

### 2018-12-29 neu_torque.tex
typos
modification v1.0

\begin{abstract}
Cosmological observations are promising ways to improve of our understandings of the neutrino mass properties. The upper bound on their sum of mass is given by the cosmic microwave background and large scale structures. These measurements are all parity-even. Here we show that, the presence of neutrino mass provides a unique contribution to the directions of the angular momentum  of galaxies, which is the first parity-odd neutrino effect of galaxies or halos. This parity-odd observable is free of the contamination of linear perturbation theory, and can be cleanly separated from other non-gravitational effects. A complete 21-cm survey deep to redshift 1 can give a $5\sigma$ confidence level of detecting the neutrino torque effect if the sum of neutrino masses is 0.05 eV.
\end{abstract}

% insert suggested PACS numbers in braces on next line
\pacs{}
% insert suggested keywords - APS authors don't need to do this
%\keywords{}

@@

\textit{Introduction.---}
Neutrino mass is a long-standing physics problem. 
The flavour oscillation experiments \citep{2002PhRvL..89a1301A} discovered the mass 
splittings of neutrinos and placed a lower bound of the sum of their mass 
$M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}. 
The existence of neutrino mass has profound impacts on cosmic evolution, and the current cosmic microwave background observations provide an 
upper bound of $M_\nu\lesssim 0.23$ eV \citep{2016A&A...594A..13P}. 
In the cosmic structure formation, neutrinos are non-relativistic, 
and contribute to the matter energy density $\Omega_m$. 
Unlike the majority of matter, the cold dark matter (CDM) and baryons, 
neutrinos maintain a high velocity dispersion, 
known as ``free-streaming'', which reduces their gravitational collapse on small scales. 
A number of large scale structure (LSS) surveys \citep{2011arXiv1110.3193L,2015AAS...22533605E}
%(such as Euclid \cite{2011arXiv1110.3193L},
%LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D}, WFIRST\cite{2012SPIE.8442E..1UG}, 
%and DESI \cite{2015AAS...22533605E})
will improve this upper bound using neutrino effects on LSS, 
however we usually encounter difficulties in disentangling neutrino effects from other parameters, 
and in understanding halo bias and cosmic variance. 
Recently, new nonlinear neutrinos effects on LSS are proposed to give independent measurements on neutrino mass \citep{2014PhRvL.113m1301Z,2016PhRvL.116n1301Z,2017NatAs...1E.143Y}. 
For these effects one needs to carefully separate and exclude the contaminations from non-gravitational contributions. 
The only clean, gravitational-only neutrino effect on LSS is the damping of weak gravitational lensing power spectrum \cite{2011arXiv1110.3193L}. 
However the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky. 
Is there a 3-dimensional measure that is cleanly gravity-driven, to measure the neutrino mass?

The angular momentum direction of galaxies. 
The direction of angular momentum (hereafter spin) is readily observable, while the magnitude is not. 
Majority of galaxies are disk galaxies, and the rotation axis is perpendicular to the disk \citep{1998MNRAS.297L..71M}. 
The orientation of the disk is obtained by inclination and parallactic angles (parity-even) and dust absorption, and from the Doppler effect of spectral lines we determine the plus-minus sign of the spin, which is parity-odd.
A parity-odd spin measurement cannot be contaminated by the linear perturbation theory. 
Only gravity affects spin statistics of halos or galaxies at large separations, since baryons do not travel to the distance of the free-streaming of neutrinos.

We present the neutrino torque effect, to modulate the spin of dark matter halos.
Here we use halos to represent galaxies.
The galaxy spin can be observed to a much larger radii through the radio emission of the gas, and the spin change only very modestly to larger radii. This suggests that the spins of a galaxy and the whole halo are well correlated.

@@

The initial halo spin is written in Lagrangian space as 
$\spin_L\propto-\int_{V_L}\bs{q}\times\bs{\nabla}\phi_c\diff^3\bs{q}$, 
where $\phi_c$ is the gravitational potential of CDM, 
$\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$. 
In the tidal torque theory \citep{1984ApJ...286...38W},
it can be written as,
$\spin_T\propto\bs{\epsilon}\,\I_q\,\T_c$, where
$\I_q=(I_{ij})\equiv(\int_{V_L} q_i q_j \diff^3\bs{q})$,
$\T_c=(T_{ij})\equiv(\partial_i\partial_j\phi_c)$
are the protogalactic inertia tensor and the local tidal shear tensor\footnote{$\I_q$ 
and $\T_c$ differ from textbook by trace, which does not contribute to $\spin_T$.}, 
and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol to collect the asymmetric components generated by
the misalignment between $\I_q$ and $\T_c$. Here, $\I_q$ and $\T_c$ are parity-even, $\bs{\epsilon}$ is parity-odd, and thus $\spin_T$ is parity-odd.
In linear perturbation theory, $\phi_c$ remains constant, so $\T_c$ decays as $a^{-2}$ due to the cosmic expansion, where $a$ is the scale factor.
%\footnote{$\phi_c(q,t)=\phi_c(q,t=0)$, $x=aq$, so $T_{x}=\partial^2_{x}\phi_c=\frac{1}{a^2}\partial^2_q\phi_c\propto\frac{1}{a^2}$.}
Also, the protohalo shrinks in size and turns more spherical in shape.
A decayed tidal field is hard to torque small and round objects, so the spin are expected to be contributed mostly in the linear regime.
These concepts can be straightforwardly tested in numerical simulations. 
In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so the spin direction is not affected by baryonic effects. 
If a halo has a merging history, it simply collects disconnected regions in Lagrangian space but conserves their total spin.

Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to halos and galaxies.
In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their free-streaming. 
They contribute a predictable tidal tensor $\T_\nu(m_{\nu_i})$ depending on the their mass \citep{2015PhRvD..92b3502I}. 
The interplay between $\I_q$ and $\T_\nu$ leads to an additional neutrino torque $\spin^T_\nu\propto\bs{\epsilon}\,\I_q\,\T_\nu$.
Integrated it over the cosmic evolution, we denote the total neutrino modulation of the halo spin as $\spin^T_{\nu 0}$ given by the tidal torque theory.
%\begin{equation}\label{eq.jnu0}
%\spin^I_{\nu 0}=\int a(\tau)D(\tau)\bs{\epsilon}\I(\tau)\T_\nu(\tau)\diff\tau.
%\end{equation}
%Here, $\tau$ is the Newtonian time, $D(\tau)$ is the growth factor, and other quantities are functions of time.

@@

\textit{Discussion.---}
We found that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than our previous understandings (e.g. \citep{2000ApJ...532L...5L}). 
We carefully investigate the numerical errors that may affect the results. 
P3M (particle-particle particle-mesh) algorithms result in higher cross-correlation between initial and final spins, compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation. 
Higher mass halos in a given simulation generally have slightly higher cross-correlations between initial and final spins, however the correlation is enhanced as we use higher mass resolutions. 
All other cross-correlation measurements have only weak dependencies on the halo mass, even in a fixed simulation. 
With different box sizes, mass resolutions, force resolutions, we find that the results are consistent across these simulations. 
The estimation of number of halo spins needed to detect the neutrino torque depends on $\mu=0.19$. This has a very weak dependence on halo mass and configuration of the simulation. An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles/fluids, further studies of reconstruction of $\I_R$ (Appendix \ref{app.I}) to calculate more precise nonlinear neutrino effects on an evolving halo. These require future simulations with computing power comparable to that of TianNu \citep{2017RAA....17...85E}.

Surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} or a modified 21cm Cosmic Vision \citep{2018arXiv180207216D} are able to observe order of $10^9$ HI galaxies ($>10^{12}M_\odot$), 
in a cosmic volume $(4{\rm Gpc}/h)^3$ below redshift $z\simeq 1$ \citep{2004MNRAS.350.1210Z}. 
Under the standard model of the Universe with $M_\nu=0.05$ eV, a $5\sigma$ detection will be reachable.
Beyond standard models, e.g., the neutrino mass could be generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}, the tidal torque history will be different, and we need different numbers of galaxies to differentiate between these models.

@@

\textit{Conclusion.---}
Enormous efforts has been contributed to the neutrino mass properties.
Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physic mysteries. 
The angular momentum is a 3-dimensional, gravity-driven, parity-odd measure, 
and is readily observable, well modeled by the tidal torque theory, and well conserved over the cosmic evolution.
It contains comparable amount of information as density field but poorly appreciated in application.
Recent developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space, 
to accurately reconstruct the neutrino torque effects,
and to unveil the neutrino mass properties in upcoming galaxy surveys.

\textit{Acknowledgments.---}
We acknowledge funding from NSERC.
The simulations were performed on the Sunnyvale cluster at CITA and on the Niagara supercomputer at the SciNet HPC Consortium.

>>>>

\begin{abstract}
Cosmological observations are promising ways to improve our understandings of the neutrino mass properties. The upper bound on their sum of mass is given by the cosmic microwave background and large scale structures. These measurements are all parity-even. Here we show that, the presence of neutrino mass provides a unique contribution to the directions of the angular momentum  of galaxies, which is the first parity-odd neutrino effect of galaxies or halos. This parity-odd observable is free of the contamination of linear perturbation theory, and can be cleanly separated from other non-gravitational effects. A complete 21-cm survey deep to redshift 1 can give a $5\sigma$ confidence level of detecting the neutrino torque effect if the sum of neutrino masses is 0.05 eV.
\end{abstract}

% insert suggested PACS numbers in braces on next line
\pacs{98.80.-k}
% insert suggested keywords - APS authors don't need to do this
%\keywords{}

@@

\textit{Introduction.---}
Neutrino mass is a long-standing physics problem. 
The flavour oscillation experiments \citep{2002PhRvL..89a1301A} discovered the mass 
splittings of neutrinos and placed a lower bound of the sum of their mass 
$M_\nu \equiv \sum_{i=1}^3 m_{\nu_i} \gtrsim$ 0.05 eV \citep{2014ChPhC..38i0001O}. 
The existence of neutrino mass has profound impacts on cosmic evolution, and the current cosmic microwave background observations provide an 
upper bound of \tcb{ $M_\nu\lesssim 0.12$ eV \citep{2018arXiv180706209P}}. 
\tcb{At low redshifts, neutrinos become non-relativistic,} 
and contribute to the matter energy density $\Omega_m$ in the structure formation. 
Unlike the majority of matter, the cold dark matter (CDM) and baryons, 
neutrinos maintain a high velocity dispersion, 
referred to as ``free-streaming'', which reduces their gravitational collapse on small scales. 
A number of large scale structure (LSS) surveys \citep{2011arXiv1110.3193L,2015AAS...22533605E}
%(such as Euclid \cite{2011arXiv1110.3193L},
%LSST \cite{2009arXiv0912.0201L}, eBOSS \cite{2016AJ....151...44D}, WFIRST\cite{2012SPIE.8442E..1UG}, 
%and DESI \cite{2015AAS...22533605E})
plan to improve this upper bound using neutrino effects on LSS,
however we usually encounter difficulties in disentangling it from other non-gravitational effects, 
and in understanding halo bias and cosmic variance. 
\tcb{Thus most popular cosmological neutrino mass measurements are not proven to work.
For example, although the neutrino effect on weak lensing power spectrum \citep{2011arXiv1110.3193L} uses pure gravity, 
there are a large set of systematics affecting the observations and unclear baryonic effects that can contaminate the neutrino signal.}
Besides, the amount of information and constraining power of the lensing is limited by the projection onto the 2-dimensional sky. 
Is there a 3-dimensional measure that is cleanly gravity-driven, to measure the neutrino mass?
%Many new neutrino effects on LSS are proposed \citep{2014PhRvL.113m1301Z, 2016PhRvL.116n1301Z, 2017NatAs...1E.143Y}.

The angular momentum direction of galaxies. 
The \tcb{\it direction} of angular momentum (hereafter \tcb{\it spin}) is readily observable, while the magnitude is not. 
Majority of galaxies are disk galaxies, and the rotation axis is perpendicular to the disk \citep{1998MNRAS.297L..71M}. 
The orientation of the disk is obtained by inclination and parallactic angles (parity-even) and dust absorption, and from the Doppler effect of spectral lines we determine the plus-minus sign of the spin, which is parity-odd.
A parity-odd spin measurement cannot be contaminated by the linear perturbation theory. 
Only gravity affects spin statistics of halos or galaxies at large separations, since baryons do not travel to the distance of the free-streaming of neutrinos.

We present the neutrino torque effect, to modulate the spin of dark matter halos.
Here we use halos to represent galaxies.
The galaxy spin can be observed to a much larger radii through the radio emission of the gas, and the spin change only very modestly to larger radii \citep{wiki:Galaxy_rotation_curve}. 
\tcb{Recent hydrodynamic galaxy formation simulations suggest that, although uncorrelated in amplitude, the spin directions of stellar, 
cold gas components and CDM halo are strongly correlated (e.g. section 4.2 of \citep{2018arXiv180407306J}). 
%In galaxy clusters, central galaxies and satellites are also found to be aligned with halos \citep{2016MNRAS.460.3772S}.
}

@@

The initial halo spin is written in Lagrangian space as 
$\spin_L\propto-\int_{V_L}\bs{q}\times\bs{\nabla}\phi_c\diff^3\bs{q}$, 
where $\phi_c$ is the gravitational potential of CDM, 
$\bs{q}$ is the Lagrangian coordinates relative to the center of mass of the protohalo in volume $V_L$. 
In the tidal torque theory \citep{1984ApJ...286...38W},
it can be written as,
$\spin_T\propto\bs{\epsilon}\,\I_q\,\T_c$, where
$\I_q=(I_{ij})\equiv(\int_{V_L} q_i q_j \diff^3\bs{q})$,
$\T_c=(T_{ij})\equiv(\partial_i\partial_j\phi_c)$
are the protogalactic inertia tensor and the local tidal shear tensor\footnote{$\I_q$ 
and $\T_c$ differ from textbook by trace, which does not contribute to $\spin_T$.}, 
and $\bs{\epsilon}=(\epsilon_{ijk})$ is the Levi-Civita symbol to collect the asymmetric components generated by
the misalignment between $\I_q$ and $\T_c$. Here, $\I_q$ and $\T_c$ are parity-even, $\bs{\epsilon}$ is parity-odd, and thus $\spin_T$ is parity-odd.
In linear perturbation theory, $\phi_c$ remains constant, so $\T_c$ decays as $a^{-2}$ due to the cosmic expansion, where $a$ is the scale factor.
%\footnote{$\phi_c(q,t)=\phi_c(q,t=0)$, $x=aq$, so $T_{x}=\partial^2_{x}\phi_c=\frac{1}{a^2}\partial^2_q\phi_c\propto\frac{1}{a^2}$.}
Also, the protohalo shrinks in size and turns more spherical in shape.
A decayed tidal field is hard to torque small and round objects, so the spin are expected to be contributed mostly in the linear regime \citep{2002MNRAS.332..325P}.
If a halo has a merging history, it simply collects disconnected regions in Lagrangian space but conserves their total spin.
These concepts can be straightforwardly tested in $N$-body simulations.

In Lagrangian space, CDM and baryonic matter are torqued by the same gravitational shear, so 
\tcb{their subsequent local evolution should not systematically change their respective spin directions.
For example, baryonic feedback has to conserve angular momentum, unless the baryons are expelled from the galaxy and carry angular momentum away systematically.
The dynamical friction between baryons and dark matter can only increase the angular momentum correlation.
These all indicate a high correlation between CDM and baryons (e.g. \citep{2018arXiv180407306J}).}

Massive neutrinos contribute sub-percent fraction of the matter ingredient, and their unique spatial distribution and evolution should contribute a unique torque to \tcb{CDM and baryons}.
In particular, neutrino density field traces CDM on large scales while their small scale structures are smoothed out by their free-streaming. 
They contribute a predictable tidal tensor $\T_\nu(m_{\nu_i})$ depending on the their mass \citep{2015PhRvD..92b3502I}. 
The interplay between $\I_q$ and $\T_\nu$ leads to an additional neutrino torque $\spin^T_\nu\propto\bs{\epsilon}\,\I_q\,\T_\nu$.
Integrated it over the cosmic evolution, we denote the total neutrino modulation of the halo spin as $\spin^T_{\nu 0}$ given by the tidal torque theory.
%\begin{equation}\label{eq.jnu0}
%\spin^I_{\nu 0}=\int a(\tau)D(\tau)\bs{\epsilon}\I(\tau)\T_\nu(\tau)\diff\tau.
%\end{equation}
%Here, $\tau$ is the Newtonian time, $D(\tau)$ is the growth factor, and other quantities are functions of time.
\tcb{
This torque is unique in that, the neutrino free-streaming scale is much larger than the scale to which local nonlinear and baryonic effects can systematically contribute.
}

@@

\textit{Discussion.---}
\tcb{
We proposed a new approach to measure the neutrino mass that has not been considered before. 
The applicability relies on the new research studies on the angular momentum connection.
Galaxy formation simulations like \cite{2018arXiv180407306J,2016MNRAS.460.3772S} are needed to study the Lagrangian spin correlations for baryons, 
despite that combining \cite{2018arXiv180407306J} and our result already implies a strong correlation. 
%For galaxy clusters, the correlation between minor axes in \cite{2016MNRAS.460.3772S} also implies the correlation between spins. After all, however, since most galaxies are not in clusters so the contribution of clusters is not a major effect.
We do not propose to use galaxies in clusters, and most spiral galaxies are not in clusters.
}

We found that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than our previous understandings (e.g. \citep{2000ApJ...532L...5L}). 
We carefully investigate the numerical errors that may affect the results. 
P3M (particle-particle particle-mesh) algorithms result in higher cross-correlation between initial and final spins, compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation. 
Higher mass halos in a given simulation generally have slightly higher cross-correlations between initial and final spins, however the correlation is enhanced as we use higher mass resolutions. 
All other cross-correlation measurements have only weak dependencies on the halo mass, even in a fixed simulation. 
With different box sizes, mass resolutions, force resolutions, we find that the results are consistent across these simulations. 
The estimation of number of halo spins needed to detect the neutrino torque depends on $\mu=0.19$. This has a very weak dependence on halo mass and configuration of the simulation. An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles/fluids, further studies of reconstruction of $\I_R$ (Appendix \ref{app.I}) to calculate more precise nonlinear neutrino effects on an evolving halo. These require future simulations with computing power comparable to that of TianNu \citep{2017NatAs...1E.143Y,2017RAA....17...85E}.

Surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} or a modified 21cm Cosmic Vision \citep{2018arXiv180207216D} are able to observe order of $10^9$ HI galaxies ($>10^{12}M_\odot$), 
in a cosmic volume $(4{\rm Gpc}/h)^3$ below redshift $z\simeq 1$ \citep{2004MNRAS.350.1210Z}. 
Under the standard model of the Universe with $M_\nu=0.05$ eV, a $5\sigma$ detection will be reachable.
Beyond standard models, e.g., the neutrino mass could be
generated through a neutrino vacuum condensate triggered by a gravitational $\theta$-term
\citep{2016PhRvD..93k3002D}, the tidal torque history will be different, 
and we need different numbers of galaxies to differentiate between these models.

@@

\textit{Conclusion.---}
Enormous efforts has been contributed to the neutrino mass properties.
Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physics mysteries. 
The angular momentum is a 3-dimensional, gravity-driven, parity-odd measure, 
and is readily observable, well modeled by the tidal torque theory, and well conserved over the cosmic evolution.
It contains comparable amount of information as density field but poorly appreciated in application.
Recent developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space, 
to accurately reconstruct the neutrino torque effects,
and to unveil the neutrino mass properties in upcoming galaxy surveys.

\textit{Acknowledgments.---}
\tcb{We thank the anonymous referee who gives valuable comments which improved this letter.} 
HRY thanks Derek Inman, Yipeng Jing, Pengjie Zhang and Jiaxin Han for helpful comments and suggestions.
We acknowledge funding from NSERC.
The simulations were performed on the Sunnyvale cluster at CITA and on the Niagara supercomputer at the SciNet HPC Consortium.
